### 2024-05-10 article.tex
final fixes done

\section{Dataset Analysis}
Two mediums vision and audio are being used for
classification of animals, two different datasets comprising of
labeled images and labeled audio are required. ImageNet was our
the initial choice for image data but due to the misrepresentation of
species, the idea was dropped, and iNaturalist was chosen as our
source for image data for its reliability and proper
representation\cite{r9}.
The image dataset has a total of 44 classes. Of the 44 classes, there are 31 mammals, 9 birds, and 4 reptiles. Moreover, we couldn’t find research-grade audio data for groups of animals other than birds. Thus, the audio model comprises 23 classes of birds all found in Nepal.
\subsection{Image Data from iNaturalist}
iNaturalist is an online social network of people sharing biodiversity information. It has a large size of data which includes images, audio, and video of large number of species found around the world. The data is collected by users who submit it to the site under the public domain, Creative Commons, or with all rights reserved. iNaturalist encourages its users to opt for more open licenses like Creative Commons for the research community. We collected a total of 38,442 images spanning 44 species all found in Nepal. As the data contained some anomalous images, we had to manually remove them. After the cleaning step, the data was preprocessed as per the Efficient Net Model for better training. Of the 38,442 images collected, 50 images from each class were used as test datasets.
API provided by iNaturalist was used for downloading the data. The bar chart in Figure \ref{image-file-count-animal} shows the frequency of each class of image.

@@

\begin{center}
	\captionof{table}{Audio Model Metrics}
	\begin{tabular}{ |c|c|c| }
		\hline
		\textbf{Parameters}      & \textbf{Metrics}          \\
		\hline
		Batch Size               & 64                        \\
		\hline
		Epochs                   & 15                        \\
		\hline
		Pretrained Model         & EfficientNetB0            \\
		\hline
		Weights                  & ImageNet                  \\
		\hline
		Learning Rate            & 0.0001                    \\
		\hline
		Weight Decay             & 0                         \\
		\hline
		Loss Function            & Categorical Cross Entropy \\
		\hline
		Optimizer                & Adam                      \\
		\hline
		Total Parameters         & 4,037,011                 \\
		\hline
		Trainable Parameters     & 4,037,011                 \\
		\hline
		Non-Trainable Parameters & 0                         \\
		\hline
	\end{tabular}
	\label{table:audio-model}
\end{center}
We see a similar pattern as seen in the vision model. Here too is the audio model with similar featured data being mislabeled. Figure \ref{accuracy-audio} and Figure \ref{loss-audio} show the accuracy and loss of the audio model respectively. Figure \ref{cm_audio} shows the confusion matrix for the audio model on the test data.
\begin{figure*}[ht]
	\centering
	\begin{minipage}[ht]{0.45\textwidth}
		\centering
		\includegraphics[width=0.8\linewidth]{Graphics/audio_accuracy.png}
		\caption{Accuracy of Audio Model on Train and Test Data}
		\label{accuracy-audio}
	\end{minipage}
	\hfill
	\begin{minipage}[ht]{0.45\textwidth}
		\centering
		\includegraphics[width=0.8\linewidth]{Graphics/audio_loss.png}
		\caption{Loss of Audio Model on Train and Test Data}
		\label{loss-audio}
	\end{minipage}
\end{figure*}

@@

\clearpage
\section{Conclusion}
Our research work in wildlife classification combines audio and visual technologies, showing promising accuracy levels of 90.39\% and 86.1\% for audio and vision models, respectively. Despite challenges in distinguishing similar species, our study represents a significant advancement in leveraging technology for wildlife conservation. By enhancing monitoring capabilities, we aim to contribute to sustainable conservation practices and promote harmonious coexistence between humans and wildlife.

>>>>

\section{Dataset Analysis}
As two mediums, vision and audio are being used for
classification of animals, two different datasets comprising of
labeled images and labeled audio are required. ImageNet was our
the initial choice for image data but due to the misrepresentation of
species, the idea was dropped, and iNaturalist was chosen as our
source for image data for its reliability and proper
representation\cite{r9}.
The image dataset has a total of 44 classes. Of the 44 classes, there are 31 mammals, 9 birds, and 4 reptiles. Moreover, we couldn’t find research-grade audio data for groups of animals other than birds. Thus, the audio model comprises 23 classes of birds all found in Nepal.
\subsection{Image Data from iNaturalist}
iNaturalist is an online social network of people sharing biodiversity information. It has a large size of data which includes images, audio, and video of large number of species found around the world. The data is collected by users who submit it to the site under the public domain, Creative Commons, or with all rights reserved. iNaturalist encourages its users to opt for more open licenses like Creative Commons for the research community. We collected a total of 38,442 images spanning 44 species all found in Nepal. As the data contained some anomalous images, we had to manually remove them. After the cleaning step, the data was preprocessed as per the Efficient Net Model for better training. Of the 38,442 images collected, 50 images from each class were used as test datasets.
API provided by iNaturalist was used for downloading the data. The bar chart in Figure \ref{image-file-count-animal} shows the frequency of each class of image.

@@

\begin{center}
	\captionof{table}{Audio Model Metrics}
	\begin{tabular}{ |c|c|c| }
		\hline
		\textbf{Parameters}      & \textbf{Metrics}          \\
		\hline
		Batch Size               & 64                        \\
		\hline
		Epochs                   & 15                        \\
		\hline
		Pretrained Model         & EfficientNetB0            \\
		\hline
		Weights                  & ImageNet                  \\
		\hline
		Learning Rate            & 0.0001                    \\
		\hline
		Weight Decay             & 0                         \\
		\hline
		Loss Function            & Categorical Cross Entropy \\
		\hline
		Optimizer                & Adam                      \\
		\hline
		Total Parameters         & 4,037,011                 \\
		\hline
		Trainable Parameters     & 4,037,011                 \\
		\hline
		Non-Trainable Parameters & 0                         \\
		\hline
	\end{tabular}
	\label{table:audio-model}
\end{center}
We see a similar pattern as seen in the vision model. Here too, the audio model is struggling to classify audio of similar species. Figure \ref{accuracy-audio} and Figure \ref{loss-audio} show the accuracy and loss of the audio model respectively. Figure \ref{cm_audio} shows the confusion matrix for the audio model on the test data.
\begin{figure*}[ht]
	\centering
	\begin{minipage}[ht]{0.45\textwidth}
		\centering
		\includegraphics[width=0.8\linewidth]{Graphics/audio_accuracy.png}
		\caption{Accuracy of Audio Model on Train and Test Data}
		\label{accuracy-audio}
	\end{minipage}
	\hfill
	\begin{minipage}[ht]{0.45\textwidth}
		\centering
		\includegraphics[width=0.8\linewidth]{Graphics/audio_loss.png}
		\caption{Loss of Audio Model on Train and Test Data}
		\label{loss-audio}
	\end{minipage}
\end{figure*}

@@

\clearpage
\section{Conclusion}
Our research work in wildlife classification combines audio and visual technologies, showing promising accuracy levels of 90.39\% and 86.1\% for audio and vision models on test data, respectively. Moreover, the audio and vision model achieves an F1 score of 86.12 and 88.12 on test data respectively. Despite challenges in distinguishing similar species, our study represents a significant advancement in leveraging technology for wildlife conservation. By enhancing monitoring capabilities, we aim to contribute to sustainable conservation practices and promote harmonious coexistence between humans and wildlife.
